---
title: 'Reproducible Research: Peer Assessment 1'
author: "Rui Mendes"
date: "January 2015"
output:
  html_document:
    keep_md: yes
---

---

#### Loading and preprocessing the data

###### Loading the csv file (using `read.csv` method) into a data.frame `activityData`.  

```{r echo=TRUE}
## Loading the data
activityData <- read.csv("data/activity.csv")

```

#### Exploring and summarizing the loaded data
See a little sample of the data and class of all data columns in order to find out what transformations need to be done.  
Also checking the number of `NAs` in the steps column
```{r echo=TRUE}
## List the first 5 elements
head(activityData, 5)

## List the last 5 elements
tail(activityData, 5)

## Checking the summary of all data columns
summary(activityData)

## Checking the class of all data columns
str(activityData)

## Checking the number of NAs in the steps column
table(is.na(activityData$steps))
```


#### Process/transform the data into a format suitable for our analysis
As we saw in the last step, the date colum must to be converted to date format
```{r echo=TRUE}
## Convert 'steps' column to numerical
steps_as_numeric <- as.numeric(activityData$steps)
activityData[,1] <- steps_as_numeric

## Convert 'date' column to Date type
date_as_date <- as.Date(activityData$date)
activityData[,2] <- date_as_date

## Convert 'interval' column to numerical
interval_as_numeric <- as.numeric(activityData$interval)
activityData[,3] <- interval_as_numeric

## Check the interval of dates of teh study
min(activityData$date)
max(activityData$date)
```



