---
title: 'Reproducible Research: Peer Assessment 1'
author: "Rui Mendes"
date: "January 2015"
output:
  html_document:
    keep_md: yes
---

---

#### Loading and preprocessing the data

###### Loading the csv file (using `read.csv` method) into a data.frame `activityData`.  

```{r echo=TRUE}
## Loading the data
activityData <- read.csv("data/activity.csv")

```

#### Exploring and summarizing the loaded data
See a little sample of the data and class of all data columns in order to find out what transformations need to be done.  
Also checking the number of `NAs` in the steps column
```{r echo=TRUE}
## List the first 5 elements
head(activityData, 5)

## List the last 5 elements
tail(activityData, 5)

## Checking the summary of all data columns
summary(activityData)

## Checking the class of all data columns
str(activityData)

## Checking the number of NAs in the steps column
table(is.na(activityData$steps))
```


#### Process/transform the data into a format suitable for our analysis
As we saw in the last step, the date colum must to be converted to date format
```{r echo=TRUE}
## Convert 'steps' column to numerical
steps_as_numeric <- as.numeric(activityData$steps)
activityData[,1] <- steps_as_numeric

## Convert 'date' column to Date type
date_as_date <- as.Date(activityData$date)
activityData[,2] <- date_as_date

## Convert 'interval' column to numerical
interval_as_numeric <- as.numeric(activityData$interval)
activityData[,3] <- interval_as_numeric

## Check the interval of dates of teh study
min(activityData$date)
max(activityData$date)
```

#### What is mean of total number of steps taken per day?
For this part of the assignment, we can ignore the missing values in the dataset.
 
 1. Make a histogram of the total number of steps taken each day
 2. Calculate and report the mean and median total number of steps taken per day
 
```{r echo=TRUE, fig.width=10}
## Aggregate number of steps per each day using sum function
stepsPerDay <- aggregate(steps ~ date, data = activityData, FUN = sum, na.action=na.omit)

## Plot the histogram of the total number of steps taken each day
barplot(stepsPerDay$steps, main="Graph of Total Steps taken per Day", names.arg = stepsPerDay$date, 
        xlab = "Days", ylab = "Total number of steps", las=3, cex.axis=0.5, cex.names=0.5)

## Calculate the mean of total number of steps taken per day
mean(stepsPerDay$steps)
## Calculate the median of total number of steps taken per day
median(stepsPerDay$steps)
```

#### What is the average daily activity pattern?
 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE, fig.width=10}
## Aggregate te number of steps per interval using the mean function
stepsByInterval <- aggregate(steps ~ interval, data = activityData, FUN = mean, na.action=na.omit)
## Make the series plot of Average Steps against Interval
plot(stepsByInterval, type = "l", main="Time-Series of Average Steps against Interval", xlab="5-minute Interval",
      ylab="Average Steps (average across all days)")

## Calculate 5-minute interval that contains the maximum number of steps
stepsByInterval$interval[which.max(stepsByInterval$steps)]
```




